---
title: "Joins"
author: "Celia Evans, PhD"
date: "March 20, 2023"
editor: visual
format:
    html:
      code-fold: true
---

## Loading Packages

```{r}
library(tidyverse)

```

## Make example dataframes

```{r}
name <- c("Alice","Bob", "Angharad","Rhys")
age <- c(23,31,25,32)
gender <- c("F","TM","NB","M")
income_range <- c("30-50K","50-75K","30-50K","100-125K")

report_2022 <- tibble(name,age,gender,income_range)

report_2022

```

```{r}
name <- c("Alice","Bob","Sally", "Angharad")
age <- c(23,31,41,25)
gender <- c("F","TM","TF","NB")
income_range <- c("50-75K","50-75K","75-100K","30-50K")

report_2023 <- tibble(name,age,gender,income_range)

report_2023
```

```{r}
income_range <- c("20-30K","30-50K","50-75K","75-100K","100-125K","125-150K")
marginal_tax_rate <- c(0.10,0.11, 0.12, 0.15,0.20,0.21)

rate_table <- tibble(income_range, marginal_tax_rate)

rate_table
```

## Keys Primary and Foreign

A **primary key** is a variable in a data frame that uniquely identifies an observation, e.g. name, SSN, Serial Number, etc.

A **foreign key** is a variable in a data frame that is the primary key in another data frame.

#### Exercise 1:

Identify the primary and foreign keys of the above dataframes.

### Left and Right Join

Think of two data frames, one we will call the **left** dataframe, the other the **right** dataframe.

An **leftjoin is a method of combining information from two data frames based on keys.**

```{r}
left_join(report_2022,report_2023, by="name")

```

```{r}
right_join(report_2022,report_2023, by="name")
```

### Exercise 2

What is the difference between the left join and the right join in the above two examples.

### Inner and full joins

An inner join only retains rows that occur in both data sets. A full join can also be called an "outer join"

```{r}
inner_join(report_2022,report_2023,by="name")
```

On the other hand:

```{r}
full_join(report_2022, report_2023, by="name")
```

### Exercise 3

a\. What is the difference between the inner join and the out join as seen before.

b\. What is the tax rate for each person in 2023? Create a dataframe giving only the name and tax rate for each person in 2023.
